Three Algorithms for Merging Navigable Graphs

### 2024-10-09 paper/main.tex
Rename parameters in pseudocode of neighborhood construction procedure
add several citations to the paper

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{url}

\title{Three Algorithms for Merging Navigable Graphs}
\author{}
\date{}

@@

\section{Introduction}
In recent years, navigable small-world graphs have become a cornerstone of efficient approximate nearest neighbor (ANN) search in high-dimensional spaces, a problem prevalent in fields such as recommendation systems, image retrieval, and natural language processing. Algorithms like Hierarchical Navigable Small World (HNSW) and its predecessor Navigable Small World (NSW) provide a robust and scalable framework for organizing data into graph structures that support fast similarity searches. However, one limitation inherent to these methods is the lack of a natural mechanism for node deletion, making the need for merging and compacting graphs crucial for maintaining the efficiency and accuracy of the search index over time.

@@

In this work, we propose three novel algorithms designed to efficiently merge navigable graphs, focusing on HNSW but extending applicability to other graph-based data structures like NSW, NSG. The merging algorithms—"Naive Merge," "Merge1," and "Merge2"—offer progressively advanced methods to combine graphs while preserving the properties that make navigable graphs effective for nearest neighbor search. These methods ensure that the resulting merged index excludes deleted nodes and maintains the structural integrity needed for fast search operations.

@@

We start from the basic search over the graph.

@@

\begin{algorithm}
\caption{\textsc{KNN-Neighborhood-Construction}$(u,U, k)$}
\label{alg:knntrategy}
\textbf{Require:} vertex $u \in V$, neighbor candidates $U \subseteq V$, $k \in \mathbb{N} $\\
\textbf{Ensure:} selected neighbors $U' \subseteq U$
\begin{algorithmic}[1]

    \State $U' \gets \text{k-closest neighbors of } U \text{ to } u$
    \State \Return $U'$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{\textsc{RNG-Neighborhood-Construction}$(u, U, m)$}
\label{alg:rngstrategy}
\textbf{Require:} vertex $u \in V$, neighbor candidates $U \subseteq V$, maximum neighborhood size $m \in \mathbb{N}$\\
\textbf{Ensure:} selected neighbors $U' \subseteq U$
\begin{algorithmic}[1]
    \State sort $v \in U$ in ascending order of $\delta(u,v)$
    \State $U' \gets \emptyset$
    \For{$v \in U$}
        \State $f \gets \text{true}$
        \For{$w \in U'$}
            \If{$\delta(u,v) \geq \delta(v,w)$}
                \State $f \gets \text{false}$
                \State \textbf{break}
            \EndIf
        \EndFor
        \If{$f$}
            \State $U' \gets U' \cup \{v\}$
        \EndIf

        \If{$|U'| \geq m $}
            \State \textbf{break}
        \EndIf
    \EndFor
    \State \Return $U'$
\end{algorithmic}
\end{algorithm}

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{authblk}

\title{Three Algorithms for Merging Navigable Graphs}
\author[1]{Alexander Ponomarenko}
\date{}

\affil[1]{YugaByte LLC}
% \affil[2]{HSE Univerity}

@@

\section{Introduction}
In recent years, navigable small-world graphs have become a cornerstone of efficient approximate nearest neighbor (ANN) search in high-dimensional spaces, a problem prevalent in fields such as recommendation systems, image retrieval, and natural language processing. Algorithms like Hierarchical Navigable Small World (HNSW) \cite{hnsw} and its predecessor Navigable Small World (NSW) \cite{nsw2011,nsw2012,nsw2014} provide a robust and scalable framework for organizing data into graph structures that support fast similarity searches. However, one limitation inherent to these methods is the lack of a natural mechanism for node deletion, making the need for merging and compacting graphs crucial for maintaining the efficiency and accuracy of the search index over time.

@@

In this work, we propose three novel algorithms designed to efficiently merge navigable graphs, focusing on HNSW but extending applicability to other graph-based data structures like NSW \cite{nsw2011}, NSG \cite{NSG}. The merging algorithms—"Naive Merge," "Merge1," and "Merge2"—offer progressively advanced methods to combine graphs while preserving the properties that make navigable graphs effective for nearest neighbor search. These methods ensure that the resulting merged index excludes deleted nodes and maintains the structural integrity needed for fast search operations.

@@

We start from the basic search over the sigle graph.

@@

\begin{algorithm}
\caption{\textsc{KNN-Neighborhood-Construction}$(v^*,C, k)$}
\label{alg:knntrategy}
\textbf{Require:} vertex $v^* \in V$, neighbor candidates $C \subseteq V$, $k \in \mathbb{N} $\\
\textbf{Ensure:} selected neighbors $C' \subseteq C$
\begin{algorithmic}[1]

    \State $C' \gets \text{k-closest neighbors of } C \text{ to } v^*$
    \State \Return $C'$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{\textsc{RNG-Neighborhood-Construction}$(v^*, C, m)$}
\label{alg:rngstrategy}
\textbf{Require:} vertex $v^* \in V$, neighbor candidates $C \subseteq V$, maximum neighborhood size $m \in \mathbb{N}$\\
\textbf{Ensure:} selected neighbors $C' \subseteq C$
\begin{algorithmic}[1]
    \State sort $v \in C$ in ascending order of $\rho(v^*,v)$
    \State $C' \gets \emptyset$
    \For{$v \in U$}
        \State $f \gets \text{true}$
        \For{$w \in C'$}
            \If{$\delta(v^*,v) \geq \delta(v,w)$}
                \State $f \gets \text{false}$
                \State \textbf{break}
            \EndIf
        \EndFor
        \If{$f$}
            \State $C' \gets C' \cup \{v\}$
        \EndIf

        \If{$|C'| \geq m $}
            \State \textbf{break}
        \EndIf
    \EndFor
    \State \Return $C'$
\end{algorithmic}
\end{algorithm}
